Synthesis of Linear Functional Programs

### 2021-07-09 article/article.tex
Write formal system start

@@

However, a programming language only with the canonical linear types lacks
expressiveness. To enhance \synname\ and its language, and consequently synth
more interesting programs -- and thus empirically prove the feasability of
more relevant synthesis in a linear context, we extend the syntax and type
system with richer types: -- polimorfic, refinements, adts parametrizáveis...  todo

@@

>>>>

Escrever introdução melhor que tenha isto como seguimento...

@@

However, a programming language only with the canonical linear types lacks
expressiveness. To enhance \synname\ and its language, synth more interesting
programs -- and thus empirically proving the feasability of more relevant
synthesis in a linear context, we extend the syntax and type system with
recursiveness and richer
types: -- polimorfic, refinements, adts parametrizáveis...  todo

@@

The development of \synname\ was initially found on an almost direct reading of
concrete inference rules and theoretical principles; and work on its environment
was also guided by formal systems and methods (e.g. the Hindley-Milner type
inference~\cite{}).  However, with the addition of recursion and richer types to the synth and
language, formal guidelines to support the implementation become scarcer, or
more complex. Accordingly, rather than derived from explicit rules,
solutions appear first to tame the extensions' inherent complex
problems (such as infinite recursion) and from them emerge implicit rules.

With this section we intend to formalize the implicit rules that guide the
synthesis of more expressive programs, alongside the already explicit rules that
model the core of the synth process, resulting in a \emph{sound} set of rules
that hopes to serve as a source of thoughts on taming synthesis of recursive
functions with richer types in a linear context, as well as describe the system
in enough detail for the synthetiser to be reproducible by a theory-driven
different implementation.

Note that \emph{sound} and \emph{complete} mean two very different
things, and that this set of rules is only sound, meaning it might generate
correct results but simultaneously hiding an infinite amount of valid programs,
and that it reflects the quite subjective trade-offs we decided on, and that
different synths might take completely different approaches.
